HO unification from object language to meta language

### 2024-03-27 paper/main.tex
update maybeeta detection

In the examples above, the first expression is a \maybeeta since $A_x$ can
reduce to $x$, the second one is not a \maybeeta since for any substitution for
$A_x$, $x$ is not free in $f \appsep x$. The third equation is a bit more
complicated since, we have a vector of lambdas, this means that the whole term
is a \maybeeta, if the inner $\lambda$-term is an $\eta$-expansion of a term
$t$, and $t$ can be reduced to a term on the form $t' x$ where $x$ is not free
in $t'$. Indeed, eq.~\ref{eq:lamlam} is a \maybeeta under the substitution 
$\sigma = \{A \mapsto \lambda x.x, B \mapsto \lambda x. \lambda y. x\}$.

@@

\begin{definition}[reduce-to]
  For any term $t$, $\lambda x_1 \ldots x_n.t_{x_0\ldots x_n}$ reduces to a bound
  variable $x$ if one of the three following cases is satisfied: 1)
  $n = 0$ and $t = x$; 2) $t$ is the application of $x$ to a list
  of term $l$ and each $l_i$ reduces to $x_i$; 3) $t$ is a variable with scope
  $s$, and for any $bv \in [x|\vec{x}]$, it exists a $s_i$, such that $s_i$
  reduces to $bv$
\end{definition}

\begin{definition}[occurs-rigidly]
  Given a term $t$, a bund variable ``rigidly'' occurs in \subterm{t} this
  term does not appear in the scope of a variable
\end{definition}

TODO: maybe eta is over-approximation

>>>>

In the examples above, the first expression is a \maybeeta since $A_x$ can
reduce to $x$, the second one is not a \maybeeta since for any substitution for
$A_x$, $x$ is not free in $f \appsep x$. The third equation is a bit more
complicated since, we have a spine of lambdas, this means that the whole term
is a \maybeeta, if the inner $\lambda$-term is an $\eta$-expansion of a term
$t$, and $t$ can be reduced to a term on the form $t' x$ where $x$ is not free
in $t'$. Indeed, eq.~\ref{eq:lamlam} is a \maybeeta under the substitution 
$\sigma = \{A \mapsto \lambda x.x, B \mapsto \lambda x. \lambda y. x\}$.

@@

\begin{definition}[reduce-to]
  A term $t$ reduce to a name $x$, if $\exists \sigma, \sigma t = x$. In
  particular, for any term $t$, $\lambda x_1 \ldots x_n.t_{x_1\ldots x_n}$
  reduces to a bound variable $x$ if one of the three following cases is
  satisfied: 1) $n = 0$ and $t = x$; 2) $t$ is the application of $x$ to a list
  of term $l$ and each $l_i$ reduces to $x_i$; 3) $t$ is a variable with scope
  $s$, and for any $bv \in [x|\vec{x}]$, it exists a $s_i$, such that $s_i$
  reduces to $bv$
\end{definition}

\begin{definition}[occurs-rigidly]
  A name $x$ occurs rigidly in a term $t$, if $\forall \sigma, \exists t' \in
  \subterm{t}, t' = x$, i.e. $x$ does not appear in the scope of any variable
  in $t$.
\end{definition}

Finally, we can derive the implementation for \maybeeta detection:

\begin{definition}[\maybeeta detection]
  For any term $t$, $\lambda x_1 \ldots x_n.t_{x_1\ldots x_n}$ is a \maybeeta if
  1) either $t$ is the application of a term with argument $l_1 \ldots l_m$ such that 
  $m \geq n$, every $l_{m-n+i}$ reduce to $x_i$, and no $l_{1\ldots m-n-1}$
  reduces to a name in $\vec{x}$; 2) or $t$ is a variable with scope $s$ and
  each $x_i$ it exists $s_j$ such that $s_j$ reduces $x_i$.
\end{definition}

As a final remark, the \maybeeta detection defined just before is an
over-approximation, in the sense that there exists some terms $t$ considered as
\maybeeta, such that forall substitution $\sigma$, $\sigma t$ is not an
$\eta$-expansion. A small example is $\lambda x. f \appsep A_x \appsep A_x$. On
the other hand, we also point out that the \maybeeta detection spot out
potential $\eta$-expansion for terms that are not in \llambda. For example,
$\lambda x.F \appsep G_x$ is considered as \maybeeta, since we have the
application of term whose argument can reduce to $x$.

The implementation we propose for the \maybeeta relation is given below.

% Here a complex maybeeta example
% \begin{gather}
%   T = \lambda y. f \appsep A_{xy} \appsep (B \appsep a \appsep (\lambda z.y\appsep C_z)) \appsep D_x
% \end{gather}
